\documentclass[showkeys,preprint]{revtex4-2}
%\documentclass[showkeys,preprint]{article}
\usepackage[utf8]{inputenc}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{lipsum}

\begin{document}
\title{A Hamiltonian theory for whistler chorus wave in the magnetosphere}
\author{Jiangshan Zheng}
\affiliation{School of Physics, Beihang University, Beijing, 100191, China}
\author{Ge Wang}
\affiliation{Institute for Fusion Studies, The University of Texas, Austin, Texas 78712, USA}
\affiliation{Center for Plasma Theory and Computation, Madison, Wisconsin 53706, USA}
\author{Bo Li}
\affiliation{School of Physics, Beihang University, Beijing, 100191, China}
\date{\today}
\date{\today}
\begin{abstract}
In this paper, we develop a novel Hamiltonian theory for the nonlinear resonant interactions between the energetic electrons and chorus wave in the magnetosphere.
Our theory studies the dynamics of the deeply trapped resonant electrons and the slowly varying coherent wave envelope in the reference frames moving with the local resonance.
We construct a new canonical transformation to separates the motion scales, and the Hamiltonian of the resonant particle is transformed to those on each local reference frames.
The Vlasov equation of the local distribution function is readily given and coupled with the slowly varying envelope equation derived from the Ampere's law self-consistently.
We further discuss the onset situation, i.e., the early chirping stage with small frequency variation, where we set the reference frame with time independent velocity at resonance of the linear most unstable wave to simplify the theoretic descriptions. 
We also consider the adiabatic regime and the water bag approximation for the evolution of resonant phase space structure.
The adiabatic regime is verified through a novel numerical simulation based on our scale separated theory.
Our theory and the corresponding numerical method are able to provide high resolution in phase space with affordable computation cost, which is urgently required to reveal the underlying physics of the nonlinear frequency chirping problem.
\end{abstract}
\maketitle

\input{pop/intro}

\input{pop/theory}

\input{pop/discussion}

\input{pop/conclusion}

\begin{acknowledgments}
    The work was inspired by H.L.Berk during one of our authors (Ge Wang) who worked with him at the University of Texas.
    As Herb's former Ph.D student and a long-time collaborator, Ge witnessed Herb's kindness, diligence, and persistence.
    Thank you, Herb.
    This work was supported by the National Natural Science Foundation of China under Grant No.~12275013.
\end{acknowledgments}

%\bibliographystyle{unsrt}
\bibliography{ref.bib}
\end{document}
