\documentclass[showkeys,preprint]{revtex4-2}
%\documentclass[showkeys,preprint]{article}
\usepackage[utf8]{inputenc}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{lipsum}
\include{math_command}

%\usepackage{xeCJK}
%\usepackage{geometry}
%\geometry{a4paper,scale=0.8}

\begin{document}
\title{A Hamiltonian theory for whistler chorus wave in the magnetosphere}
\author{Jiangshan Zheng}
\author{Ge Wang}
\author{Bo Li}
\affiliation{School of Physics, Beihang University, Beijing, 100191, China}
\date{\today}
\begin{abstract}
    In this paper, we develop a novel Hamiltonian theory for the nonlinear resonant interactions between the energetic electrons and chorus wave in the magnetosphere.
    Our theory studies the dynamics of the deeply trapped resonant electrons and the slowly varying coherent wave envelope in the reference frames moving with the local resonance.
    We construct a new canonical transformation to separates the motion scales, and the Hamiltonian of the resonant particle is transformed to those on each local reference frames.
    The Vlasov equation of the local distribution function is readily given and coupled with the slowly varying envelope equation derived from the Ampere's law self-consistently.
    We further discuss the onset situation, in which we consider the early chirping stage with small frequency variation.
    For the onset stage, we can set the reference frame with the resonance velocity of the linear most unstable wave, with time independent, thus simplifies the theoretic descriptions. 
    We also consider the adiabatic regime and the water bag approximation for the evolution of resonant phase space structure, it demonstrates the chirping behavior of the chorus wave packet.
    Based on our scale separated theoretic description, novel numerical simulation method can be readily constructed. It is able to provide high resolution in phase space with affordable computation cost, which is urgently required to reveal the underlying physics of the nonlinear frequency chirping problem.
\end{abstract}
\maketitle

\input{pop/intro}

\input{pop/theory}

\input{pop/discussion}

\input{pop/conclusion}

%\bibliographystyle{unsrt}
\bibliography{ref.bib}
\end{document}
